---
title: "Togga FPL Introduction"
author: "Jeremy Elster"
date: "September 17, 2015"
output: html_document
runtime: shiny
---


# Introduction
This year, my friends and I made the leap from the official Barclay’s Premier League Fantasy game to [Togga](http://playtogga.com). We looked forward to Togga’s draft-based version as a way to keep engagement and interpersonal interaction high. Another reason I pushed to make the swap is to gain access to Togga’s Opta-based scoring. I figured that even if my team crashed and burned, I’d be able to sift through the wealth of match and performance data to test hypotheses about the game I love.

For handling and manipulating data, I use Excel (to help pull the stats from Togga’s website) and R. The plan is to one day publish content through a R Shiny web app but for now I will be using R Markdown. Let’s get down to business!

# Togga Statistical Categories and Scoring

## Togga Statistical Categories
```{r, echo=FALSE, message=FALSE}
list.of.packages <- c("dplyr", "ggvis", "DT", "data.table")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

library(dplyr)
library(ggvis)
library(DT)
library(data.table)


```

```{r}
print(getwd())
print(cat(ls()))
print(cd("resource_folder"))
```